---
title: "assignment 5"
author: "Zongyan Wang"
date: "Friday, November 27, 2015"
output: pdf_document
---
# Assignment 5
## Remove the previous file
```{r}
outputs <- c(# downloadData.R
  "gapminder.tsv",                     
              # summaryPlot and convertDataFormat.R
          "gapminder.csv", "gdpPercapOverCountry.png", "popOverCountry.png",
          "lifeExpOverCountry.png", "gdpPercapPlot.png", "popPlot.png",
          "lifeExpPlot.png",
          # linearRegression.R 
      "lifeExpOnYear.csv", "Top3EconomylifeExp.png")
file.remove(outputs)
```
## download file to current folder
This step requires package "downloader" to solve the rmarkdown process may occur.
```{r}
source("downloadData.R")
```
## Read that data from "downloadData.R", perform some analysis and write numerical data to file in CSV
```{r, warning = FALSE}
source("summaryPlot and convertDataFormat.R")
```
## Read the data converted by "summaryPlot and convertDataFormat.R", fit linear regression and save the result to folder.
```{r, warning = FALSE}
source("linearRegression.R")
```
Please visit my Github page: https://github.com/wangzongyan for more R and Phython code.